---
title: "Functionalizing NEON Rasters"
author: "Enrique Monta&ntilde;o"
date: "June 21, 2016"
output: html_document
---

## Objectives

Objectives/tasks to work though:

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, set values == 0 to NA (not trees)
1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize density and plot vertical cutoff lines.
1. Plot the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

## Load Libraries
```{r load-libraries}
library(raster)
library(rgdal)
library(rhdf5)

```

## Import CHM Raster

```{r import-chm}
source("scripts/functions.r")

# setting some parameters
# raster filename
chm_file <- "../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif"
# import the raster
chm <- raster(chm_file)

chosen_breaks <- c(0,6,30,100)

# set no data values; trees are > 0
chm[chm == 0] <- NA

#calculate reclass matrix
rcl_m <- create_height_class_matrix(chosen_breaks)

# plot density
plot_chm_density(chm, site.name="Teakettle", bins = chosen_breaks)

# export result to pdf
make_pdf(plot_chm_density(rast_in = chm,
                          site.name = "Teakettle",
                          bins = chosen_breaks),
         filename = paste("TEAK_chm_density.pdf"),
         height=6, width=7)

# Perform the reclass
chm_rcl <- reclassify(chm,
                      rcl_m)


plot_reclassified_raster(rast.in = chm, site.name = "Teakettle", breaks = chosen_breaks)

make_pdf(plot_reclassified_raster(rast.in = chm, site.name = "Teakettle", breaks = chosen_breaks),
         filename = "TEAK_reclassified_chm.pdf",
         height=7, width=6)

```

## Export the result

```{r export-chm}
# export geotiff
# writeRaster(chm_rcl,
#             filename="reclassify_Teak_chm.tif",
#             format="GTiff",
#             options="COMPRESS=LZW",
#             overwrite = TRUE,
#             NAflag = -9999)

```

## Document the environment

```{r doc-logfile}
# sink(paste0(format(Sys.time(), "%y-%m-%d:%H%M%S"),
#      "_sessionInfo.txt"))
# sessionInfo()
# sink()
```




